Better fusion for filters

### 2014-04-04 2014betterfusionforfilters/section/01-Introduction.tex
explicitly state similarities and differences to existing ILP formulations

\begin{itemize}
\item   We present an algorithm to cluster the operators of a data flow graph into sub-graphs that can be fused into individual imperative loops. We express the constraints on fusion as an integer linear program, and use an off-the-shelf solver to compute a valid clustering. \TODO{ref}

>>>>

\begin{itemize}
\item   
We use existing integer linear programming clustering formulations such as Megiddo\cite{megiddo1998optimal} and Darte\cite{darte2002contraction},
and modify them to express constraints on fusing loops of different sizes if they can be fused into generator loops of the same size.
\TODO{ref}
% We present an algorithm to cluster the operators of a data flow graph into sub-graphs that can be fused into individual imperative loops.
% We express the constraints on fusion as an integer linear program, and use an off-the-shelf solver to compute a valid clustering. \TODO{ref}

\item
We present a simplification to constraint generation that would also be applicable to existing integer linear programming formulations,
where constraints between two nodes need not be generated if there exists a fusion-preventing path between the two. \TODO{ref}

### 2014-04-04 2014betterfusionforfilters/section/04-ILP.tex
explicitly state similarities and differences to existing ILP formulations

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (an edge from $i$ to $j$)            \\
           \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (no edge from $i$ to $j$)            \\
           \> \ldots
\end{tabbing}
Here, we look at every pair of nodes $i$ and $j$.

@@

\paragraph{Fusion-preventing edges}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (fusion-preventing edges from $i$ to $j$)      \\
           \> \ldots
\end{tabbing}
For every fusion-preventing edge, we add the constraint $x_{ij} = 1$, so that no fusion can occur.
This, together with the precedence constraints above, has the effect of enforcing $\pi_i < \pi_j$.

\paragraph{Fusion between different types}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\emptyset \wedge gen_j=\emptyset$)                 \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\{i'\}$) \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_j=\{j'\}$) \\
           \> \ldots
\end{tabbing}
Nodes with different loop sizes cannot be fused together, generally.
If two nodes have different sizes or rates, but are both sub-rates of another rate, they \emph{can} be fused together if they are both fused with their generators.
\TODO{Expand}.

\paragraph{Array contraction}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (for all edges from i)            \\
           \> \ldots \\
\end{tabbing}
An array is fully contracted if all its output edges are fused with it.
Thus, $c_i=0$ only if $\forall j | (i,j) \in E. x_{ij} = 0$.
By minimising $c_i$ in the objective function, we favour solutions that reduce the number of required intermediate arrays.

>>>>

\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \>    $x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (an edge from $i$ to $j$)            \\
           \> $-N x_{ij}$ \> $\le$ \> $\pi_j - \pi_i$ \> $\le$ \> $N x_{ij}$ \\
           \>             (no edge from $i$ to $j$)            \\
           \> \ldots
\end{tabbing}
As per Megiddo\cite{megiddo1998optimal}, we look at every pair of nodes $i$ and $j$.

@@

\paragraph{Fusion-preventing edges}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (fusion-preventing edges from $i$ to $j$)      \\
           \> \ldots
\end{tabbing}
As per Megiddo\cite{megiddo1998optimal}, for every fusion-preventing edge, we add the constraint $x_{ij} = 1$, so that no fusion can occur.
This, together with the precedence constraints above, has the effect of enforcing $\pi_i < \pi_j$.

\paragraph{Fusion between different types}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $=$   \> $1$             \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\emptyset \wedge gen_j=\emptyset$)                 \\
           \> $x_{i'j}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_i=\{i'\}$) \\
           \> $x_{ij'}$   \> $\le$ \> $x_{ij}$        \>       \>            \\
           \> (if $\tau_i \not= \tau_j \wedge gen_j=\{j'\}$) \\
           \> \ldots
\end{tabbing}
The main difference to existing integer linear programming solutions is that, while nodes with different loop sizes cannot generally be fused together, we do support some fusion of filtered sizes.
If two nodes have different sizes or rates, but are both sub-rates of another rate, they \emph{can} be fused together if they are both fused with their generators.
\TODO{Expand}.

\paragraph{Array contraction}
\begin{tabbing}
MMMMM   \= MMM \= M \= MMM \= M \= MMM \= \kill
Minimise   \> \ldots \\
Subject to \> \ldots \\
           \> $x_{ij}$    \> $\le$ \> $c_i$           \>       \>            \\
           \> (for all edges from i)            \\
           \> \ldots \\
\end{tabbing}
As per Darte's work on optimising for array contraction\cite{darte2002contraction}, we define a variable $c_i$ for each array.
An array is fully contracted if all its output edges are fused with it.
Thus, $c_i=0$ only if $\forall j | (i,j) \in E. x_{ij} = 0$.
By minimising $c_i$ in the objective function, we favour solutions that reduce the number of required intermediate arrays.
